\documentclass[a4paper,12pt]{report}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{multicol}
\usepackage{lscape}

\newcommand{\gvmtic}{\textsc{gvmtic}}
\newcommand{\gvmtc}{\textsc{gvmtc}}
\newcommand{\glink}{\textsc{gvmtlink}}
\newcommand{\gvmtas}{\textsc{gvmtas}}
\newcommand{\gvmtcc}{\textsc{gvmtcc}}
\newcommand{\gvmtxc}{\textsc{gvmtxc}}

% Title Page
\title{The Glasgow Virtual Machine Toolkit Manual}
\author{Mark Shannon}
\date{ } 

\begin{document}
\maketitle 

\begin{abstract}
The GVMT is a toolkit for building virtual machines. The GVMT does not dictate the overall design of the virtual machine, but it does eliminate a lot of implementation details. 

The developer needs to develop an interpreter and supoorting code, as normal. The toolkit handles garbage collection transparently
and can automatically create a just-in-time compiler from the interpreter source code.

This manual covers the components of the Glasgow Virtual Machine Toolkit, hereafter called the GVMT, and their use. 

Convertion of source-code to bytecodes is not handled by the GVMT, that is up to the developer.

\tableofcontents

\end{abstract}


\chapter{Introduction}

\input{./intro.tex}

\chapter{The tools\label{chap:tools}}

\input{./tools.tex}

\chapter{GVMT Execution Model\label{chap:abstract}}

\input{./model.tex}

\chapter{Building a VM using the Toolkit}

\section{Before you start}
Before starting a virtual machine, you will need to generate some bytecode programs to execute. If you are targetting a pre-existing VM format, this is straightforward. If you are creating a new language you will need a parser. \gvmtic{} can produce a C header file containing definitions for all the opcodes: the C macro for the opcode for instruction `xxx' of interprter 'int' is \verb|GVMT_OPCODE(int, xxx)|.

\section{Defining the components}
For GVMT to build a VM, you will need to define:
\begin{itemize}
\item An interpreter. See section \ref{sect:gvmtic}
\item Any other support code that your interpreter calls. See section \ref{sect:gvmtc}
\end{itemize}
You will also need to define a small number of functions required by the GVMT. The header file for these can be found in /include/user.h and sample implementations can be found in \verb|/example/native.c| and \verb|/example/support.c|
It may also be useful to specify:
\begin{itemize}
\item The object layout for heap objects. See section \ref{sect:objects}
\end{itemize}

Once these are defined, the VM is built as follows, see figure \ref{fig:build}: The \verb|.gsc| files for the interpreter and other support code is created by running \gvmtic{} and \gvmtc{} respectively. These  \verb|.gsc| files are converted to \verb|.gso| with \gvmtas. All the  \verb|.gso| files are then linked together using \glink{} to form as single object file, which can be linked with any native object files to form an executable using the standard linker.

\section{Debugging}
Although the toolkit maintains symbolic information from the source to the executable, some variables may get renamed. 
The signatures of functions may appear different in the debugging, although their names should be unaltered.
The execution stack can be accessed via the stack pointer, the name of which is \verb|gvmt_sp|.

\section{Adding a compiler}
It is probably best to have a working interpreter before adding the compiler. To generate the compiler simply run \gvmtcc{} on the output file from \gvmtic. The resulting C++ file should be compiled and linked using the system C++ compiler. See section \ref{sect:gvmtcc} for more details.

\chapter{User interface}
\input{./user.tex}

\appendix

\chapter{Installing the GVMT}

\input{./install.tex}

\chapter{GSC Instruction set\label{app:inst}}
\small

\input{./instructions.tex}

\end{document}
